---
title: "Data Envelopment Analysis"
subtitle: "Adapted Example from Lu Hao (Georgetown) Using Public Data"
author: "Jason P. Casey"
date: 2019-04-01
output: html_notebook
---

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="user-functions.R")

# Aaditional libraries
library(rDEA)

set.seed(1965)

# constants and other presets

# new names for IPEDS file
new_names <- c("unitid", "Institution Name", "Collection Year", "Carnegie Classification", "State", "Control", "Land Grant",
               "Full-time Enrollment", "Part-time Enrollment", "Student FTE",
               "Associates Degrees", "Bachelors Degrees", "Masters Degrees", 
               "Research Doctorate Degrees", "Professional Doctorate Degrees", "Other Doctorate Degrees",
               "Six-year Graduation Rate",
               "Instruction per FTE (GASB)", "Research per FTE (GASB)", "Public Support per FTE (GASB)",
               "Academic Support per FTE (GASB)", "Student Services per FTE (GASB)", 
               "Institutional Support per FTE (GASB)", "Other Exp per FTE (GASB)",
               "Instruction per FTE (FASB)", "Research per FTE (FASB)", "Public Support per FTE (FASB)",
               "Academic Support per FTE (FASB)", "Student Services per FTE (FASB)",
               "Institutional Support per FTE (FASB)", "Other Exp per FTE (FASB)", 
               "Total Staff FTE", "Faculty FTE", "Instructional Staff FTE")
```

# Read Data

## IPEDS Excerpt

The following except was pulled by creating a custom data file on the [IPEDS](https://nces.ed.gov/ipeds/use-the-data) data page at NCES.  (My kingdom for an API to call!)  It has the usual verbose column names which I've replaced with less cryptic labels.  These data are from 2017-18, except finance, which lags a year behind.

```{r read_ipeds}
ipeds <- read_csv("data/ipeds_extract.csv",
                  col_names = new_names,
                  skip = 1) %>%
  filter(Control %in% c("Private not-for-profit", "Public")) %>%
  mutate(`Instruction Expenses per FTE` = ifelse(Control == "Public",
                                                 `Instruction per FTE (GASB)`,
                                                 `Instruction per FTE (FASB)`),
         `Research Expenses per FTE` = ifelse(Control == "Public",
                                              `Research per FTE (GASB)`,
                                              `Research per FTE (FASB)`),
         `Public Service Expenses per FTE` = ifelse(Control == "Public",
                                                    `Public Support per FTE (GASB)`,
                                                    `Public Support per FTE (FASB)`),
         `Academic Support per FTE` = ifelse(Control == "Public",
                                             `Academic Support per FTE (GASB)`,
                                             `Academic Support per FTE (FASB)`),
         `Student Services Expenses per FTE` = ifelse(Control == "Public",
                                                      `Student Services per FTE (GASB)`,
                                                      `Student Services per FTE (FASB)`),
         `Institutional Support Expenses per FTE` = ifelse(Control == "Public",
                                                           `Institutional Support per FTE (GASB)`,
                                                           `Institutional Support per FTE (FASB)`),
         `Other Expenses per FTE` = ifelse(Control == "Public",
                                           `Other Exp per FTE (GASB)`,
                                           `Other Exp per FTE (FASB)`),
         `Total Enrollment` = `Full-time Enrollment` + `Part-time Enrollment`,
         `Weighted Graduation Efficiency` = ((`Associates Degrees` + `Bachelors Degrees`) +
                  1.5 * `Masters Degrees` +
                  2 * (`Research Doctorate Degrees` +
                         `Professional Doctorate Degrees` +
                         `Other Doctorate Degrees`)) /
           `Total Enrollment`,
         `Full-time Administrative FTE` = `Total Staff FTE` - `Faculty FTE`,
         `Administrative Expenses` = (`Academic Support per FTE` +
                                        `Student Services Expenses per FTE`+
                                        `Institutional Support Expenses per FTE` +
                                        `Other Expenses per FTE`) * `Student FTE`,
         `Student Expenses` = (`Instruction Expenses per FTE` +
                                 `Student Services Expenses per FTE`) * `Student FTE`) %>%
  select(-(`Instruction per FTE (GASB)`:`Other Exp per FTE (FASB)`))

ipeds
```

## THE Data

This extract was pulled for the 2019 publication of the [Times Higher Education World University Rankings](https://www.timeshighereducation.com/world-university-rankings/2019/world-ranking#!/page/0/length/25/sort_by/rank/sort_order/asc/cols/stats).  It was pulled using a screen-scraping script in Python.  These data align with the 2017-18 IPEDS extract.  Ranks and scores were estimated using the underlying scores using the methodology published in the guide.  _All estimated ranks and scores matched those published for the top 100 institutions._  These data were used to pull the citations and research scores.

```{r read_the}
the_data <- read_csv("data/the_data.csv") %>%
  rename(inst_rank = rank)

the_data %>%
  ggplot(aes(x = score, y = inst_rank)) +
  geom_point(color = "darkblue",
             alpha = 0.15)
```

## Combine Data

```{r}
dea_data <- ipeds %>%
  inner_join(the_data, by = "unitid")

dea_data %>%
  ggplot(aes(fte_enroll, `Student FTE`)) +
  geom_point(color = "darkblue",
             alpha = 0.1)
```

```{r}
models <- tribble(
  ~model, ~in.vars, ~out.vars,
  'Instruction', c('StudentRelatedExpdPerFTE', 'FTEFaculty', 'Faculty_Std_Ratio'), c('6Yr_GradRate', 'TotalDegrees', 'WGE_Score'),
  'Research', c('FTEFaculty', 'ResearchExpdperFctFTE'),c('ResearchScore_THE','CitationScore_THE', 'GovGrantsContracts'),
  'Administrative Cost',c('TotalAdminCost', 'FullTimeStaff'),c('TotalDegrees','FTEStudentEnrollment')
)

dea_stat <- function(model) {
  # feasible
  # thetaOpt
  # lambda
  # lambda_sum
  
  df <- model %>%
    mutate(feasible = map(1),
           theta = map(2),
           lambda = map(3),
           lambda_sum = map(4))
  
  return(df)
}

# insts <- select(institutions, SchoolName_code)

models <- models %>%
  mutate(inputs = map(in.vars, grab_frame, original = institutions),
         outputs = map(out.vars, grab_frame, original = institutions),
         dea.model = map2(inputs, outputs, run_dea),
         feasible = map(dea.model, 1),
         theta = map(dea.model, 2),
         # theta = map2(insts, theta, tibble),
         lambda = map(dea.model, 3),
         lambda_sum = map(dea.model, 4)) %>%
  select(-in.vars, -out.vars, -dea.model)

results <- institutions %>%
  select(SchoolName_code)

for (i in seq(1:3)) {
  results <- cbind(results, models$theta[[i]])
}

names(results) = c('SchoolName_code', models$model)

```

